Parametric controls on vegetation responses to biogeochemical forcing in the CLM5

### 2019-07-01 main_revision.tex

\section{Results}
Our analysis used four sites, with eight parameters perturbed to four different values, each of those with control, CO$_{2}$ and N fertilization simulations.  We examine the parametric responses of five output variables chosen for their importance in the overall biogeochemical system (GPP, NPP, LAI, leaf nitrogen  ($N_{area}$, gN m$^{-2}$), and total vegetation carbon (g m$^{-2}$). We did not look at soil output variables as our parameter analysis is focused on vegetation responses.

@@

\subsection{Overall system state}
The parameter responses of the site-level water and nitrogen limitation (see below) in the control (unfertilized) parameter ensemble are shown in figure \ref{btran state}. GPP, NPP, LAI, leaf N content per unit area and total vegetation carbon are shown for the Niwot Ridge, (Figure \ref{NR1 state}), Caxiuan\~a (Figure \ref{CAX state}), Harvard Forest (Figure \ref{HVF state}) and Metolius (Figure \ref{MET state}) sites. In each of these figures, the y-axis corresponds to the output variable and the x-axis represents the range of parameter perturbation with -1 to +1 corresponding to the range of variation in each parameter, depicted in Table \ref{table_ranges}. Lines of differing colors illustrate responses to variation in the thirteen target parameters, with the central crossover point in each figure corresponding to the default case.

\subsubsection{Water and N limitation across sites}
The four sites varied in the degree to which they are limited by water and N availability in the control state, with important impacts upon the parametric response characteristics (Figure \ref{btran state}). Water limitation in CLM5 can be expressed via the $B_{tran,mn}$ parameter, which is a model property calculated as a function of the prognostic leaf water potential. $B_{tran}$  is used to reduce assimilation and thus stomatal conductance by scaling the photosynthetic capacity, $V_{c,max}$. Leaf water potential and  $B_{tran}$ vary diurnally, so we use the average daily minimum value here as a metric of drought stress (1 is no drought stress, 0 is total stomatal closure). For N limitation, the percentage of total NPP (where NPP is GPP-$r_{auto}$) that is used to pay for nitrogen uptake (PNU) is a direct indicator of how N stress affects the carbon budget.

@@

Thus, the sites chosen span a range of initial limiting conditions.  
Differences in the initial conditions at the sites have implications for the parametric responses. Underlying limitations by various boundary conditions (water, nutrients, light, temperature) can affect both the ability of the system to respond to fertilization and/or parameter modifications affecting other resources, as we describe further in the following sections.

\subsubsection{Parametric control over system state}
At all sites, specific leaf area (\emph{slatop}) has a large first-order impact on LAI, and on leaf N concentrations (Figure \ref{NR1 state}), as expected, since it is used directly in the calculation of both of these quantities. The impact of SLA on GPP is always positive, despite the thinner leaves having lower N, and thus lower photosynthetic capacity, per unit area of leaf. The impact of SLA on NPP can, however, sometimes be negative (CAX, NWR), since the fixed-fraction allometry model in CLM5 can generate shaded leaves that are in negative carbon balance at high LAI.

Root:leaf ratio (\emph{froot\_leaf}) has 1) a direct effect on LAI, via changes in tissue allocation fractions , 2) an indirect effect on NPP, via changes in maintenance costs of fine root pools, 3) an effect on the uptake of water via root biomass and 4) an effect on nutrient limitation through root-mediated N uptake costs.  The first two impacts decrease NPP with higher \emph{froot\_leaf} costs and the latter two increase it via \emph{froot\_leaf} benefits. CAX and HVF exhibit higher LAI with fewer roots.   NWR has large reduction in LAI (and GPP & NPP) with fewer roots. MET exhibits a mixed response crossing an optimal value. \emph{froot\_leaf} is a primary control over drought stress at both NWR and MET ( Figure \ref{btran state}).

Low values of \emph{stem\_leaf} ratio resulted in higher LAI at all sites, as well as declines in total vegetation carbon. At NWR, there is an optimum (or local maximum) stem carbon observed, due to the impacts of a too-high leaf area index, which caused declines in NPP at low \emph{stem\_leaf}. At all other sites, the impact of increasing leaf allocation is to increase NPP (almost doubling at Metolius). Decreases in stem:leaf ratio also caused a decline in leaf N per unit area, potentially on account of the higher demand for N uptake and thus higher N limitation. This effect is llustrated clearly for Niwot Ridge and Metolius in Figure \ref{btran state}).

The first-order impact of increasing \emph{N\_{costs} was to increase the respiration losses associated with nitrogen uptake.  Across the range tested here (bearing in mind that these parameters are somewhat unconstrained) there was limited impact of \emph{N\_{costs} on the system state with the exception of the Niwot Ridge site, where N limitation was highest and which showed a decline in LAI, NPP, GPP and vegetation carbon when costs were high, as well as slight increases in leaf N content where costs were low. Other sites showed more muted responses in the vegetation state.

Increasing \emph{fracfixers} affects the maximum amount of carbon that plants can pay towards symbiotic nitrogen uptake (N fixation). For all sites and variables here, it has a relatively moderate and positive impact on the control state, in particular at the Niwot and Harvard forest sites. As is the case for \emph{N\_{costs}, with high fixing capacity, N uptake expenditure is lower, N uptake high, and NPP moderately increased, resulting in increases in LAI and a slight response of GPP to that LAI increase.

Changing \emph{leafcn} -the target leaf C:N ratio- over one standard deviation from the mean TRY values, had substantial impacts on all the output variables at all sites. Higher \emph{leafcn} values decrease $N_{area}$, decreasing photosynthetic capacity and hence GPP.  In particular, the higher \emph{leafcn} values had $N_{leaf}$ and photosynthetic capacities so low that GPP was almost zero at Metolius and Harvard Forest and more than halved at Niwot Ridge (\ref{NR1 state}).  We do not find much evidence of saturation in the GPP vs. \emph{leafcn} relationships, in that GPP response does not level off at high leaf N content. This is consistent with the analysis by \cite{lawrence2018} who find that V${_c,max}$ numbers in CLM5 might be low compared to observations.  For NPP,  decreases in respiration rates caused by high \emph{leafcn}, and reduced carbon expenditure on N uptake (Figure \ref{btran state}) somewhat ameliorated the GPP increases, particularly at Caxiuan\~a, which has high ambient temperatures and respiration rates as a result. At Niwot, an optimum \emph{leafcn} ratio (or a local maximum in the parameteric response) was apparent. For NPP this was around the default value, but for LAI it was 0.5 standard deviations lower, reflecting the complex interactions between the expenditure on N uptake and acquisition of N for tissue growth.

Changes in \emph{medlyn\_slope}, the stomatal slope parameter, most directly impact GPP, but can also affect leaf temperature via changes in canopy resistance and thus sensible heat flux. Water limitation is also modified as a result of altered transpiration. At Caxiuan\~a (Figure \ref{CAX state}), the very lowest values of  \emph{medlyn\_slope} caused a complete death of the ecosystem. In Figure \ref{btran state} we can see that low values of the Medlyn stomatal slope ameliorated plant drought stress (by reducing stomatally-mediated water losses), presumably at the cost of limiting photosynthesis past the point at which positive plant growth could be maintained. At Niwot, a clear local maximum was indicated in the response to the Medlyn parameter, indicating that above this level, stomata lose too much water and below it, they do not photosynthesize enough. At Metolius and Harvard forest, only the lower \emph{medlyn\_slope} values affected the photosynthetic rates, indicating co-limitation of GPP by processes other than stomatal conductance.

@@

The \emph{N\_{costs} parameterization primarily impacts the respiration expended on N uptake and the availability of carbon for tissue growth thus, its impact is greatest for LAI (Figure \ref{LAI CO2 and N respones 2001}). At Caxiuan\~a, Niwot and Harvard forest, high  \emph{N\_{costs} resulted in stronger LAI responses to N fertilization, since these cases had higher N limitation in the first instance (Figure \ref{btran state}).  These high \emph{N\_{costs} scenarios also showed a slight increase in the CO$_{2}$ fertilization response, perhaps since their high N limitation was alleviated by extra assimilated carbon.   Low values of \emph{N\_{costs} were associated with low CO$_{2}$ responses (and vice versa) in particular at Caxiuana and Harvard Forest, which can be interpreted as a reflection of the lower C expenditure on N uptake in the control simulations.  While nitrogen uptake costs are important to the fertilization responses, they do not affect the total amount of N in the system, and therefore appeared to play a secondary role to the amount of N fixation available.

@@

There was limited impact of \emph{cn\_flex\_a} on the fertilization responses, as per it's impact on the state.  At Harvard Forest, low values of  \emph{cn\_flex\_b} increased responses to N deposition and decreased response to CO$_{2}$. Leaf N concentration followed suit, increasing substantially in the N deposition experiment. Low values of \emph{cn\_flex\_b} decrease the amount of carbon spent on N uptake when \emph{N\_costs} are high, therefore potentially suppressing leaf N content. Adding N overcomes this suppression, leading to greater fertilization responses.

@@

\subsection{The status of the CLM5 representation of plant carbon and nitrogen representation}
Our efforts to improve the fidelity of the nitrogen cycle in the CLM5 were primarily motivated by the need to move the model behaviour closer to several key dynamic features of the system. Notably, the dynamics of leaf nitrogen content with respect to environmental scarcity (\cite{lovelock1998}, \cite{brzostek2014}), the dynamics of photosynthetic capacity with respect to environmental conditions (\cite{xu2012}, \cite{ali2016},\cite{rogers2017}, \cite{bloomfield2018}), the changes in N fixation rate with vegetation state and the scarcity of N (\cite{vitousek2002}), and to include a means of deploying carbon as a means to acquire additional nitrogen (\cite{terrer2018}). The latter acts as a proxy for the numerous ways in which plant symbiont systems might adapt to low nutrient conditions, including increases in fine root exudation and allocation, fungal biomass, tissue metabolic rates and turnover/nutrient foraging.  In doing so, we integrate a suite of physiological processes that lack complete mechanistic and quantitative understanding.  By investigating the critical roles of these processes in the model dynamics, we hope to highlight a research need, and to stimulate research on and improvements in the representation of these processes.

>>>>

\section{Results}
Our analysis used four sites, with eight parameters perturbed to four different values, each of those with control, CO$_{2}$ and N fertilization simulations.  We examine the parametric responses of five output variables chosen for their importance in the overall biogeochemical system (GPP, NPP, LAI, leaf nitrogen per unit area ($N_{leaf}$, gN m$^{-2}$), and total vegetation carbon (g m$^{-2}$). We did not look at soil output variables as our parameter analysis is focused on vegetation responses.

@@

\subsection{Overall system state}
The parameter responses of the site-level water and nitrogen limitation (see below) in the control (unfertilized) parameter ensemble are shown in figure \ref{btran state}. GPP, NPP, LAI, $N_{leaf}$ and total vegetation carbon are shown for the Niwot Ridge, (Figure \ref{NR1 state}), Caxiuan\~a (Figure \ref{CAX state}), Harvard Forest (Figure \ref{HVF state}) and Metolius (Figure \ref{MET state}) sites. In each of these figures, the y-axis corresponds to the output variable and the x-axis represents the range of parameter perturbation with -1 to +1 corresponding to the range of variation in each parameter, depicted in Table \ref{table_ranges}. Lines of differing colors illustrate responses to variation in the thirteen target parameters, with the central crossover point in each figure corresponding to the default case.

\subsubsection{Water and N limitation across sites}
The four sites varied in the degree to which they are limited by water and N availability in the control state (Figure \ref{btran state}). Water limitation in CLM5 can be expressed via a model property termed $B_{tran}$, which is calculated as a function of the (prognostic) leaf water potential. $B_{tran}$  is used to reduce assimilation and thus stomatal conductance by scaling the photosynthetic capacity, $V_{c,max}$. Leaf water potential and $B_{tran}$ vary diurnally, so we use the average daily minimum ($B_{tran,mn}$) here as a metric of drought stress (1 is no drought stress, 0 is total stomatal closure). For N limitation, the percentage of total NPP (where NPP is GPP-$r_{auto}$) that is used to pay for nitrogen uptake (PNU) is a direct indicator of how N stress affects the carbon budget.

@@

Thus, the sites chosen span a range of initial limiting conditions. Underlying limitations by different boundary conditions (water, nutrients, light, temperature) affect the ability of the system to respond to fertilization and/or parameter modifications, as we describe further in the following sections.

\subsubsection{Parametric control over system state}
In this section, we discuss the impacts of the parameter variation on the initial model state.  Many parameters have both first-order direct impacts, and then further impacts via other parts of the model. Here we attempt to illustrate how these mechanisms operate in CLM5 within the parameter space explored here.

At all sites, specific leaf area (\emph{slatop}) has a large first-order impact on LAI, and on $N_{leaf}$ (Figure \ref{NR1 state}), as expected, since it is used directly in the calculation of both of these quantities. The impact of SLA on GPP is always positive, despite the thinner leaves having lower N, and thus lower photosynthetic capacity, per unit area of leaf. The impact of SLA on NPP can, however, sometimes be negative (CAX, NWR), since the fixed-fraction allometry model in CLM5 can generate shaded leaves that are in negative carbon balance at high LAI.

The root:leaf ratio (\emph{froot\_leaf}) has numerous impacts, including 1) on LAI, via tissue allocation fractions , 2) on NPP, via changes in maintenance costs of fine root pools, 3) on the uptake of water (via root length density) and 4) on nutrient limitation through root-mediated N uptake costs. The first two impacts represent the costs of more roots, and the latter two their benefits. CAX and HVF exhibit higher LAI with fewer roots. NWR has large reduction in LAI (and GPP \& NPP) with fewer roots. MET exhibits a mixed response crossing an optimal value. \emph{froot\_leaf} is a primary control over drought stress at both NWR and MET ( Figure \ref{btran state}) indicating the impact of mechanism '3' here, for very low root densities.

Low values of \emph{stem\_leaf} ratio resulted in higher LAI at all sites, as well as declines in total vegetation carbon. At NWR, the parameter range crosses a local maximum for vegetation carbon, since increases in NPP are non-linear with respect to leaf investment, resulting in lower vegetation biomass overall. At all other sites, the impact of increasing leaf allocation is to invoke large increases in LAI and NPP. Lower stem allocation and higher LAI increase the N demand, reducing $N_{leaf}$ and increase PNU at Niwot Ridge and Metolius in Figure \ref{btran state}).

Increasing $N_{costs}$ increases the respiration losses associated with nitrogen uptake.  All sites showed a moderate decline in LAI, NPP, GPP and vegetation carbon when costs were high, but lesser impacts when costs were reduced (Figures \ref{NR1 state}, \ref{CAX state} \ref{HVF state}, \ref{MET state}). In contrast to the allocation parameters, $N_{costs}$ does not appear to be a first-order control over the pre-fertilization system state.

Increasing \emph{fracfixers} affects the maximum amount of carbon that plants can pay towards symbiotic nitrogen uptake (N fixation). For all sites and variables here, it has a relatively moderate and positive impact on the control state. As is the case for $N_{costs}$, \emph{fracfixers} does not appear to be a first-order control over the pre-fertilization system state.

Changing \emph{leafcn} over one standard deviation from the mean TRY values, had substantial impacts on all the output variables at all sites. Higher \emph{leafcn} values decrease $N_{leaf}$, decreasing photosynthetic capacity and GPP.  In particular, at MET and HVF the higher \emph{leafcn} values had $N_{leaf}$ and photosynthetic capacities so low that GPP was almost zero Figures \ref{HVF state}, \ref{MET state}).  GPP continues to increase almost linearly with declining \emph{leafcn} (increasing $N_{leaf}$), indicating a lack of saturation of photosynthesis with respect to photosynthetic capacity.  This is consistent with the analysis by \cite{lawrence2018} who find that V${_c,max}$ numbers in CLM5 might be low compared to observations.  For NPP, respiration costs of modified $N_{leaf}$ and altered PNU (Figure \ref{btran state}) buffered the larger changes in GPP. At CAX, for example, the lowest \emph{leafcn} had a GPP 225 gC/m^{2}/y higher, while the corresponding NPP increase was only 52 gC/m^{2}/y.

Changes in \emph{medlyn\_slope}, the stomatal slope parameter, across the range shown here, have limited impacts on the pre-fertilized system state, with the exception of NWR, where lower values of the slope parameter generated higher photosynthesis, consistent with this site experiencing severe drought stress and thus benefiting from increased water efficiency.

most directly impact GPP, but can also affect leaf temperature via changes in canopy resistance and thus sensible heat flux. Water limitation is also modified as a result of altered transpiration. At Caxiuan\~a (Figure \ref{CAX state}), the very lowest values of  \emph{medlyn\_slope} caused a complete death of the ecosystem. In Figure \ref{btran state} we can see that low values of the Medlyn stomatal slope ameliorated plant drought stress (by reducing stomatally-mediated water losses), presumably at the cost of limiting photosynthesis past the point at which positive plant growth could be maintained. At Niwot, a clear local maximum was indicated in the response to the Medlyn parameter, indicating that above this level, stomata lose too much water and below it, they do not photosynthesize enough. At Metolius and Harvard forest, only the lower \emph{medlyn\_slope} values affected the photosynthetic rates, indicating co-limitation of GPP by processes other than stomatal conductance.

@@

The $N_{costs}$ parameterization primarily impacts the respiration expended on N uptake and the availability of carbon for tissue growth thus, its impact is greatest for LAI (Figure \ref{LAI CO2 and N respones 2001}). At Caxiuan\~a, Niwot and Harvard forest, high  $N_{costs}$ resulted in stronger LAI responses to N fertilization, since these cases had higher N limitation in the first instance (Figure \ref{btran state}).  These high $N_{costs}$ scenarios also showed a slight increase in the CO$_{2}$ fertilization response, perhaps since their high N limitation was alleviated by extra assimilated carbon.   Low values of $N_{costs}$ were associated with low CO$_{2}$ responses (and vice versa) in particular at Caxiuana and Harvard Forest, which can be interpreted as a reflection of the lower C expenditure on N uptake in the control simulations.  While nitrogen uptake costs are important to the fertilization responses, they do not affect the total amount of N in the system, and therefore appeared to play a secondary role to the amount of N fixation available.

@@

There was limited impact of \emph{cn\_flex\_a} on the fertilization responses, as per it's impact on the state.  At Harvard Forest, low values of  \emph{cn\_flex\_b} increased responses to N deposition and decreased response to CO$_{2}$. $N_{leaf}$ followed suit, increasing substantially in the N deposition experiment. Low values of \emph{cn\_flex\_b} decrease the amount of carbon spent on N uptake when N costs are high, therefore potentially suppressing $N_{leaf}$. Adding N overcomes this suppression, leading to greater fertilization responses.

@@

\subsection{The status of the CLM5 representation of plant carbon and nitrogen representation}
Our efforts to improve the fidelity of the nitrogen cycle in the CLM5 were primarily motivated by the need to move the model behaviour closer to several key dynamic features of the system. Notably, the dynamics of $N_{leaf}$ with respect to environmental scarcity (\cite{lovelock1998}, \cite{brzostek2014}), the dynamics of photosynthetic capacity with respect to environmental conditions (\cite{xu2012}, \cite{ali2016},\cite{rogers2017}, \cite{bloomfield2018}), the changes in N fixation rate with vegetation state and the scarcity of N (\cite{vitousek2002}), and to include a means of deploying carbon as a means to acquire additional nitrogen (\cite{terrer2018}). The latter acts as a proxy for the numerous ways in which plant symbiont systems might adapt to low nutrient conditions, including increases in fine root exudation and allocation, fungal biomass, tissue metabolic rates and turnover/nutrient foraging.  In doing so, we integrate a suite of physiological processes that lack complete mechanistic and quantitative understanding.  By investigating the critical roles of these processes in the model dynamics, we hope to highlight a research need, and to stimulate research on and improvements in the representation of these processes.
